```{r}
library(dplyr)
library(tidyverse)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
Sample_EU_Superstore <- read_excel("Sample - EU Superstore.xlsx")
View(Sample_EU_Superstore)
```

```{r df_eu_storeData}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
df_eu_storeData <- Sample_EU_Superstore
view(df_eu_storeData)
```
# what does each row represent?
```{r}
head(df_eu_storeData)
```
it seems like a unique combination of order id and product id

## confirming
```{r}
df_eu_storeData |> 
    distinct(`Order ID`, `Product ID`) |> 
    nrow()
```
found 9 rows with duplication

## extracting product id duplicates
```{r}
df_eu_storeData |> 
    group_by(`Order ID`) |> 
    count(`Product ID`) |> 
    head(10)
```
### another way to do the same
```{r}
df_eu_storeData |> 
    group_by(`Order ID`, `Product ID`) |> 
    summarise(n = n()) |>
    head(10)
```
## observing duplicate rows
```{r}
df_eu_storeData |> 
    group_by(`Order ID`) |> 
    count(`Product ID`) |> 
    arrange(desc(n)) |> 
    head(10)
```

```{r}
df_eu_storeData |> 
    group_by(`Order ID`) |> 
    count(`Product ID`) |> 
    filter(n > 1) |> 
    inner_join(df_eu_storeData)
```
```{r}
# Convert date columns to date format
library(lubridate)
df_eu_storeData$`Order Date` <- ymd(df_eu_storeData$`Order Date`)
df_eu_storeData$`Ship Date` <- ymd(df_eu_storeData$`Ship Date`) 
```
##Calculate necessary metrics
```{r}
library(dplyr)
df_eu_storeData |>
  group_by(`Customer ID`) |>
  summarise(date_since_first_visit = as.numeric(as.POSIXct("2019-01-31") - as.POSIXct(min(`Order Date`)), units = "days"), 
            first_transaction_date = min(`Order Date`),
            last_transaction_date = max(`Order Date`),
            total_transactions = n_distinct(`Order ID`),
            total_revenue = sum(Sales)
         ) |>
  mutate(current_date = as.Date("2019-01-31"),
        days_since_last_transaction = as.numeric(as.POSIXct(current_date) - as.POSIXct(last_transaction_date), units = "days"),median_days_between_orders = median(date_since_first_visit, na.rm = TRUE)) -> df_metric_1
          
```
##Calculate RFM metrics
```{r}
library(Hmisc)

df_rfm <- df_metric_1 |>
  mutate(interval_recency = ntile(days_since_last_transaction, 5),
         interval_frequency = ntile(total_transactions, 5),
         interval_monetary = ntile(total_revenue, 5)) |>
  mutate(R_score = dense_rank(interval_recency),
         F_score = dense_rank(interval_frequency),
         M_score = dense_rank(interval_monetary),
         RFM_score = paste0(R_score, F_score, M_score),
         RFM_segment = case_when(
           RFM_score %in% c("111", "112", "121", "131, 141", "151", "132", "123","122", "212", "211", "115", "113", "114", "132", "213") ~ "High Value Customer",
           RFM_score %in% c("135", "134", "143", "145","221", "222", "223", "232, 233", "251", "241","322", "231", "332") ~ "Loyal Customer",
           RFM_score %in% c("152", "153", "224", "225", "234", "235", "242", "243", "252", "253", "244", "245", "254", "255", "215", "214", "144","311", "411", "313", "314", "315", "312", "321", "331", "324", "334", "343") ~ "Mid Value Customer",
           RFM_score %in% c("434", "534", "535", "415", "414", "413", "424", "412", "422", "421","511", "512", "425", "513", "514", "515", "521", "522", "523", "524", "525", "353", "423", "432", "433", "355", "344", "345", "354", "355", "435", "444", "431") ~ "New Customer",
           RFM_score %in% c("441", "442", "451", "452", "531","532", "533", "542", "541", "552","551", "553", "543", "445", "454", "545", "544", "554", "555") ~ "Churn Customer",
           TRUE ~ "Can't lose" 
         )) 
saveRDS(df_rfm, "df_rfm")
```
##Visulization
```{r}
library(ggplot2)
# Bar chart
ggplot(df_rfm, aes(x = RFM_segment, fill = RFM_segment)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x = "segment", y = "No. of Customer", title = "customer Metrics Segment")

```
##category_of_products_mostly_purchased
```{r}
library(dplyr)
library(ggplot2)

category_of_products_mostly_purchased <- df_eu_storeData |>
  group_by(Category) |>
  summarise(total_quantity = sum(Quantity)) |>
  arrange(desc(total_quantity)) |>
  slice(1) |>
  pull(Category)

ggplot(df_eu_storeData, aes(x = Category, fill = Category)) +
  geom_bar() +
  xlab("Category of Products") +
  ylab("Number of Customers") +
  ggtitle("Most Purchased Category of Products") +
  geom_text(stat='count', aes(label=after_stat(count)), vjust=-0.5, size=5) +
  theme(plot.title = element_text(hjust = 0.5))

```
##highest_profit_category
```{r}
options(scipen = 999)
highest_profit_category <- df_eu_storeData |>
  group_by(Category) |>
  summarise(total_profit = sum(Profit)) |>
  arrange(desc(total_profit)) |>
  slice(1)|> 
  pull(Category)
ggplot(df_eu_storeData, aes(x = Category, y = Profit, fill = Category)) +
 geom_bar(stat = "identity") +
  xlab("Category of Products") +
  ylab("Profit") +
  ggtitle("Highest Profit Category of Products") + 
  theme(plot.title = element_text(hjust = 0.1))
```
##most_discounted_category
```{r}
library(scales)

# calculate most discounted category
most_discounted_category <- df_eu_storeData |>
  filter(!is.na(Discount)) |>
  group_by(Category) |>
  summarise(Discount = mean(Discount)) |>
  arrange(desc(Discount)) |>
  slice(1) |>
  pull(Category)
# plot mean discount by category
df_eu_storeData |>
  filter(!is.na(Discount)) |>
  group_by(Category) |>
  summarise(mean_discount = mean(Discount)) |>
  ggplot(aes(x = Category, y = mean_discount, fill = Category)) +
  geom_bar(stat = "identity") +
  scale_y_continuous(labels = percent_format()) +
  labs(x = "Category", y = "Mean Discount", fill = "Category") +
  ggtitle("Discount by Category") +
  theme_minimal()

```
# Calculate most frequent customer segment
```{r}
most_frequent_segment <- df_eu_storeData |>
  group_by(Segment) %>%
  summarise(frequency = n_distinct(`Order ID`))
  arrange(desc()) |>
  slice(1) |>
  pull(Segment)
ggplot(df_eu_storeData, aes(x = segment, fill = Segment)) +
  geom_bar() +
  xlab("Customer Segment") +
  ylab("Frequency") +
  ggtitle("Most Frequent Customer Segment") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_text(stat = "count", aes(label = after_stat(count)), 
            vjust = -0.5, size = 5) 
```
# Find the most preferred shipment mode
```{r}
df_most_preferred_shipment <- df_eu_storeData |>
  group_by(`Ship Mode`) %>%
  summarize(total_shipments = n_distinct(`Order ID`)) |>
  arrange(desc(total_shipments))

ggplot(df_most_preferred_shipment, aes(x = `Ship Mode`, fill = `Ship Mode`, y = total_shipments)) +
  geom_bar(stat = "identity") +
  labs(x = "Ship Mode", y = "Total Shipments", 
       title = "Most Preferred Shipment Mode") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_text(aes(label = total_shipments), 
            vjust = -0.5, size = 3)

```
